### 2023-06-25 SpringSummerSeverity/talk/FigureGenerator.Rmd

---
title: Figures for NAPC talk
output:
  html_document:
    keep_md: false
    theme: united
    highlight: tango
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, results='hide'}
knitr::opts_chunk$set(message = FALSE, warning=FALSE, 
                      echo=FALSE, eval=TRUE, fig.path='C:/Users/devan.mcgranahan/GithubProjects/GreatPlainsFire/SpringSummerSeverity/talk/figs/', dev=c('png', 'pdf')) 
# Packages 
  pacman::p_load(tidyverse, sf, grid, gridExtra)
  pacman::p_load_gh('devanmcg/wesanderson')
# Data 
  # regional map data
    load('C:/Users/devan.mcgranahan/GithubProjects/SpatialFireBehavior/paper/figures/mapping.Rdata')
  # study data
    cgrec_gpkg = 'C:/Users/devan.mcgranahan/GithubProjects/GreatPlainsFire/SpringSummerSeverity/gis/boundaries/CGREC_PBG_26914.gpkg'
```

```{r region_map, fig.width=8, eval = FALSE}
mapping$nd_l3 <- mapping$nd_l3 %>% st_simplify(dTolerance = 1000)
mapping$gp_l1 <- mapping$gp_l1 %>% st_simplify(dTolerance = 10000)

region_map <-
  ggplot() + theme_void(24)  +
  geom_sf(data = mapping$us, 
          fill = 'lightgrey', 
          color = NA)+ 
    geom_sf(data = mapping$gp_l1, 
            fill = wes_palettes$DeadLiveLight[1]) +
    geom_sf(data = mapping$nd,
            fill = wes_palettes$DeadLiveLight[2], 
            color = "darkgrey") +
    geom_sf(data = mapping$us, 
            fill = NA, 
            color = "white")

state_map <-
  ggplot() + theme_void(30)  +
    geom_sf(data = mapping$nd,
            fill = 'NA', 
            color = "darkgrey") + 
    geom_sf(data = filter(mapping$nd_l3,
                          L3 != 'Northwestern Great Plains'), 
            fill = wes_palettes$DeadLiveLight[1], 
            show.legend = FALSE) +
    geom_sf_label(data = filter(mapping$nd_l3,
                                L3 != 'Northwestern Great Plains'),
                  aes(label = L3), 
                  size = 6) +
    geom_sf(data = filter(mapping$pts, 
                          name != "Hettinger REC"), 
            aes(shape = feature), 
            size = 4,
            show.legend = FALSE) +
    scale_shape_manual(values = c(16, 16, 17)) +
    geom_sf_text(data = filter(mapping$pts, 
                               name != "Hettinger REC"), 
                 aes(label = name), 
                  size = 6, 
                 fontface = c("bold", 'bold', 'italic'),
                 nudge_y = c(20000, 20000, -20000), 
                 nudge_x = c(-35000, 0, -15000),
            show.legend = FALSE)

v1<-viewport(width = 1, height = 1, x = 0.5, y = 0.5) #plot area for the main map
v2<-viewport(width = 0.45, height = 0.45, x = 0.71, y = 0.75) #plot area for the inset map
print(state_map,vp=v1) 
print(region_map,vp=v2)
```

```{r BarkerBurns}
  # Fire mapping

 # st_layers(cgrec_gpkg )

  pastures <- st_read(cgrec_gpkg, 'Pastures') 
  patches <- st_read(cgrec_gpkg, 'PasturePatches')

  fires <- st_read(cgrec_gpkg, 'FirePerimeters') %>%
              mutate(Year = as.factor(Year))

    ggplot() + theme_void() +
      geom_sf(data = filter(fires, unit == 'Barker'), 
      aes(fill = Year, 
          alpha = Season)) +
      geom_sf(data = filter(patches, Unit == 'Barker'), 
              fill = NA, 
              color = 'black', 
              lty = 2, 
              size = 1) + 
      geom_sf(data = filter(pastures, Unit == 'Barker'), 
              fill = NA, 
              color = 'black', 
              lty = 1, 
              size = 2) +
      geom_sf_text(data = filter(fires, unit == 'Barker'), 
                   aes(label = Year), 
                   nudge_y = 50) +
      geom_sf_text(data = filter(fires, unit == 'Barker'), 
                   aes(label = substr(Season, 1, 3)), 
                   nudge_y = -50) +
      scale_alpha_manual(values = c(0.25, 0.8)) + 
      theme(legend.position = 'none')
```