### 2024-12-11 SpringSummerSeverity/paper/AllScriptTogetherSI.Rmd
draft SpringSummerSeverity

---
title: Evaluating an attempt to restore summer fire in the Northern Great Plains
author:
- Devan Allen McGranahan and Jay P. Angerer
subtitle:  'Supplementary Information'
header-includes:
  - \usepackage{amsmath}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage[colorlinks=true, citecolor=blue, urlcolor=blue,linkcolor=blue, pdfborder={0 0 0}]{hyperref}
  - \urlstyle{same}
  - \usepackage{xcolor} 
  - \usepackage{booktabs}
  - \raggedright
  - \usepackage{listings}
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
    toc: false
---

This document provides script for fetching remotely-sensed imagery and processing results in the \textsf{R} statistical environment.

# Remote sensing

## Sentinel-2

This \texttt{EvalScript} can be used in the [Copernicus browser](https://browser.dataspace.copernicus.eu/) as a Custom script to create a Custom visualization that can be exported as a 16-bit \texttt{.tiff} for raster analysis. 
The user must select the area of interest and imagery dates and is responsible for assessing cloudiness.

\begin{lstlisting}[language=Java]

//VERSION=3
function setup() {
  return {
    input: ["B04", "B08", "B8A", "B11", "B12"],
    output: { bands: 2, sampleType: "UINT16" }
  };
}

function evaluatePixel(sample) {
  let nbr = index(sample.B08, sample.B12);
  let ndvi = index(sample.B08, sample.B04);
  // apply offset for UINT16 
  return [10000 * nbr + 10000, 
          10000 * ndvi + 10000]; 
}
\end{lstlisting}

\clearpage
## Landsat

This script for [Google Earth Engine](https://code.earthengine.google.com/) exports a composited NDVI \texttt{.tiff} file to the user's Google Drive for the area of interest defined as \texttt{geometry} for an eight week period beginning with each date given in \texttt{listDate}. 
The script combines imagery from Landsat missions 5, 7, \& 8 and handles cloud masking.

\begin{lstlisting}[language=Java]
// Geometry
var geometry: Polygon, 4 vertices
  type: Polygon
  coordinates: List (1 element)
    0: List (5 elements)
      0: [-99.51276195869765,46.71741699210676]
      1: [-99.42521465645156,46.71741699210676]
      2: [-99.42521465645156,46.778346255312734]
      3: [-99.51276195869765,46.778346255312734]
      4: [-99.51276195869765,46.71741699210676]
  geodesic: false

// Main script

var batch = require('users/fitoprincipe/geetools:batch');

///****variables that need to be changed by user
//****ADD google drive directory name that you want to download files to
var gdrivedir = 'landsat'

var listDate= ["1992-04-20", "1992-07-20", "1993-04-20", "1993-07-20", 
               "1994-04-20", "1994-07-20", "1995-04-20", "1995-07-20", 
               "1996-04-20", "1996-07-20", "1997-04-20", "1997-07-20", 
               "1998-04-20", "1998-07-20", "1999-04-20", "1999-07-20", 
               "2000-04-20", "2000-07-20", "2001-04-20", "2001-07-20", 
               "2002-04-20", "2002-07-20", "2003-04-20", "2003-07-20",
               "2004-04-20", "2004-07-20", "2005-04-20", "2005-07-20", 
               "2006-04-20", "2006-07-20", "2007-04-20", "2007-07-20", 
               "2008-04-20", "2008-07-20", "2009-04-20", "2009-07-20",
               "2010-04-20", "2010-07-20", "2011-04-20", "2011-07-20", 
               "2012-04-20", "2012-07-20", "2013-04-20", "2013-07-20", 
               "2014-04-20", "2014-07-20", "2015-04-20", "2015-07-20",
               "2016-04-20", "2016-07-20", "2017-04-20", "2017-07-20", 
               "2018-04-20", "2018-07-20", "2019-04-20", "2019-07-20", 
               "2020-04-20", "2020-07-20", "2021-04-20", "2021-07-20",
               "2022-04-20", "2022-07-20"]

listDate.forEach(function (listDate) {
  //yearRanges.forEach(function (yearRange) {
    exportTimeseries(listDate)
  })
//})
function exportTimeseries(listDate) {

  // Defines a base date/time for the following examples.
  var startDate = ee.Date(listDate);
  print(startDate, 'The start date/time');
  print(startDate.format('YYYYMMdd'))
  var sdate = startDate.format('YYYYMMdd')
  var endDate = startDate.advance(8, 'week')
  print(endDate, 'The end date/time');
  var edate = endDate.format('YYYYMMdd')

  print(endDate.format('YYYYMMdd'))

  //from https://developers.google.com/earth-engine/tutorials/
  //                community/extract-raster-values-for-points
  //function to mask cloud and shadow pixels
  function fmask(img) {
    var cloudShadowBitMask = 1 << 4;
    var cloudsBitMask = 1 << 3;
    var qa = img.select('QA_PIXEL');
    var mask = qa.bitwiseAnd(cloudShadowBitMask).eq(0)
      .and(qa.bitwiseAnd(cloudsBitMask).eq(0));
    return img.updateMask(mask);
  }

  // Selects and renames bands of interest for Landsat OLI.
  function renameOli(img) {
    return img.select(
      ['SR_B2', 'SR_B3', 'SR_B4', 'SR_B5', 'SR_B6', 'SR_B7'],
      ['Blue', 'Green', 'Red', 'NIR', 'SWIR1', 'SWIR2']);
  }

  // Selects and renames bands of interest for TM/ETM+.
  function renameEtm(img) {
    return img.select(
      ['SR_B1', 'SR_B2', 'SR_B3', 'SR_B4', 'SR_B5', 'SR_B7'],
      ['Blue', 'Green', 'Red', 'NIR', 'SWIR1', 'SWIR2']);
  }

  // Prepares (cloud masks and renames) OLI images.
  function prepOli(img) {
    img = fmask(img);
    //img = lcfmask(img)
    img = renameOli(img);
    return img;
  }

  // Prepares (cloud masks and renames) TM/ETM+ images.
  function prepEtm(img) {
    img = fmask(img);
    //img = lcfmask(img)
    img = renameEtm(img);
    return img;
  }

  // Apply scaling factors
  function applyScaleFactors(image) {
    var opticalBands = image.select('SR_B.').multiply(0.0000275).add(-0.2);
    var thermalBands = image.select('ST_B.*').multiply(0.00341802).add(149.0);
    return image.addBands(opticalBands, null, true)
                .addBands(thermalBands, null, true);
  }

  // Get surface reflectance collections for Landsat
  // maps scaling factors and cloud masking functions

  var oliCol = ee.ImageCollection('LANDSAT/LC08/C02/T1_L2')
              .filter(ee.Filter.bounds(geometry))
              .map(applyScaleFactors)
              .map(prepOli);

  var etmCol = ee.ImageCollection('LANDSAT/LE07/C02/T1_L2')
                .filter(ee.Filter.bounds(geometry))
                .map(applyScaleFactors)
                .map(prepEtm);

  var tmCol = ee.ImageCollection('LANDSAT/LT05/C02/T1_L2')
              .filter(ee.Filter.bounds(geometry))
              .map(applyScaleFactors)
              .map(prepEtm) ;

  var landsatCol = oliCol.merge(etmCol).merge(tmCol)
                   .filterDate(startDate, endDate) ;

  // Calculate NDVI 
  var addIndices = function(image) {
     var ndvi = image.normalizedDifference(['NIR', 'Red'])
                .rename('ndvi').float();
  return image.addBands([ndvi]);
  };

  var indices = landsatCol.map(addIndices) 
                .select('ndvi');
  // Create composite of images from date range
    var composite = indices.mean() ;

    var fileName = ee.Date(startDate) 
                  .format('yyyy-MM-dd')
                  .getInfo() ;

  Export.image.toDrive({
      image: composite,
      description: fileName,
      scale: 30,
      folder: 'landsat', 
      region: geometry
      });
}
\end{lstlisting}

\newpage

# Analysis

Because the analysis script assumes hundreds of imagery files and a large Excel file with 42 years of hourly weather observations have been saved locally, it cannot be run directly from this document but is provided here for transparency and reference.

## Remotely sensed imagery

\begin{lstlisting}[language=R]

# Load necessary packages
# Note that package terra is required but is not loaded
# (called directly to not create conflicts with dplyr verbs)
  pacman::p_load(tidyverse, sf, stars, foreach, doSNOW)

# Load location boundaries (data not provided)
  cgrec_gpkg = './CGREC_PBG_26914.gpkg'

  pastures <- st_read(cgrec_gpkg, 'Pastures') 
  patches <- st_read(cgrec_gpkg, 'PasturePatches') 
#
# Get veg data for unburned pastures
#
  NoFirePts <- st_read(cgrec_gpkg, 'SamplePoints') %>%
                filter(location == 'Refuge')
  # Map to directory with Sentinel-2 imagery
    imagery_dir = 'C:/Path/To/Sentinel'
    images <- list.files(imagery_dir)
  # Use parallel processing to chug imagery
  { 
    begin = Sys.time() 
    pacman::p_load()
    cores = parallel::detectCores()
    cl <- makeCluster(cores, methods = F, useXDR = F)
    registerDoSNOW(cl)
    NoFireIndices <-
      foreach(i=1:length(images), 
              .combine = 'bind_rows',
              .errorhandling = 'remove', 
              .packages=c('tidyverse', 'sf')) %dopar% {
        image = images[i]
        image_path = paste0(imagery_dir, '/', image)
        ras <- terra::rast(image_path)
        names(ras) <- c('nbr', 'ndvi') 
        ras <- ras[['ndvi']]
        float = (ras-10000)/10000
        terra::extract(float,
          NoFirePts %>%
            select(pasture, sample) %>%
            terra::vect() , 
          FUN = mean, 
          bind = TRUE)  %>%
        st_as_sf() %>%
          as_tibble() %>%
        mutate(ImageDate = substr(image, 1, 10)) %>%
        select(ImageDate, pasture, sample, ndvi) 
            }
    stopCluster(cl)
  Sys.time() - begin 
  }
#
# Get fuel greenness and dNBR for completed burns
#
  fires <- st_read(cgrec_gpkg, 'FirePerimeters') %>%
              mutate(Year = as.factor(Year)) %>%
              filter(status == 'Completed') %>%
              unite('fire', c(unit, Pasture, Patch), sep = "-") %>%
              select(fire, Year, Season, PreBurn, PostBurn )
  SamplePts <- 
    st_read(cgrec_gpkg, 'SamplePointsRegular')
{ 
  begin = Sys.time() 
  pacman::p_load(foreach, doSNOW)
  cores = parallel::detectCores()
  cl <- makeCluster(cores , methods = F, useXDR = F)
  registerDoSNOW(cl)
  BurnIndices <-
    foreach(i=1:length(fires$fire), 
            .combine = 'bind_rows',
            .errorhandling = 'remove', 
            .packages=c('tidyverse', 'sf')) %dopar% {
              # Get fire
                fire = slice(fires, i)
              # Get sample points
                pts <-
                  fire %>%
                    st_intersection(SamplePts)
              # Get dates
                pre_date = fire$PreBurn
                post_date = fire$PostBurn
              # Fetch & process multi-band rasters
                # pre image
                  pre_image = images[substr(images, 1, 10) == pre_date]
                  pre_path = paste0(imagery_dir, '/', pre_image)
                  pre_ras <- terra::rast(pre_path) %>%
                                terra::crop(terra::vect(fire))
                  names(pre_ras) <- c('nbr', 'ndvi') 
                  pre_ras <-  (pre_ras-10000)/10000
                # post-fire
                  post_image = images[substr(images, 1, 10) == post_date]
                  post_path = paste0(imagery_dir, '/', post_image)
                  post_ras <- terra::rast(post_path)[[1]] %>%
                                terra::crop(terra::vect(fire))
                  post_ras = (post_ras-10000)/10000
                # Calculate dNBR & replace in pre-fire raster
                  d_ras = pre_ras['nbr'] - post_ras
                  names(d_ras) <- 'dNBR'
                  pre_ras[[1]] <- d_ras
              # Sample rasters
                 terra::extract( pre_ras,
                          pts %>%
                            select(-PostBurn) %>%
                            terra::vect() , 
                          FUN = mean, 
                          bind = TRUE)  %>%
                  st_as_sf() %>%
                    as_tibble() %>%
                    select(-geometry) 
            }
  stopCluster(cl)
  Sys.time() - begin 
}
#
# Get historical Landsat data for unburned pastures
#
  imagery_dir = 'C:/Path/To/Landsat'
  images <- list.files(imagery_dir)
  { 
    begin = Sys.time() 
    pacman::p_load(foreach, doSNOW)
    cores = parallel::detectCores()
    cl <- makeCluster(cores, methods = F, useXDR = F)
    registerDoSNOW(cl)
    NoFireTrends <-
      foreach(i=1:length(images), 
              .combine = 'bind_rows',
              .errorhandling = 'remove', 
              .packages=c('tidyverse', 'sf')) %dopar% {
                image = images[i]
                image_path = paste0(imagery_dir, '/', image)
                ras <- terra::rast(image_path)
                ras <- ras[['ndvi']]
                 terra::extract(ras,
                               NoFirePts %>%
                                 select(pasture, sample) %>%
                                 st_transform(4326) %>%
                                 terra::vect() , 
                               FUN = mean, 
                               bind = TRUE)  %>%
                  st_as_sf() %>%
                  as_tibble() %>%
                  mutate(ImageDate = tools::file_path_sans_ext(image))  %>%
                  select(ImageDate, pasture, sample, ndvi) 
              }
    stopCluster(cl)
    Sys.time() - begin 
  }
\end{lstlisting}

## Weather data

This script wrangles the weather data downloaded from the [North Dakota Ag Weather Network's Streeter station](https://ndawn.ndsu.nodak.edu/station-info.html?station=48).

\begin{lstlisting}[language=R]

pacman::p_load(tidyverse, readxl)

seasons <- tibble(season = c('spring', 'summer'), 
                 start = c('04-20', '07-20'), 
                 end = c('06-10', '09-10') ) %>%
          mutate(start = as.Date(start, '%m-%d'), 
                 end = as.Date(end, '%m-%d'))

BurnDays <- 
  read_xlsx('./data/BurnData.xlsx', 'BurnDays') %>%
  filter(is.na(certainty)) %>%
    mutate(date = paste(year, date), 
           date = as.Date(date, '%Y %B %d')) %>%
    select(date) %>%
  distinct()

WxData <- lst() 
# Get daily data 
  DailyWx <- 
    read_xlsx('./data/CGREC_weather.xlsx', 'daily')
# Identify rainy days
  WxData$Rainfall <-
    DailyWx %>%
    select(Year, Month, Day, Rainfall) %>% 
    unite( c(Month, Day), col = 'day', sep = '-', remove = F) %>%
    mutate(day = as.Date(day, '%m-%d')  , 
           season = case_when(
             between(day, seasons$start[1], seasons$end[1]) ~'Spring', 
             between(day, seasons$start[2], seasons$end[2]) ~'Summer',
             TRUE ~ NA      
           ) ) %>% 
     unite( c(Year, Month, Day), col = 'date', sep = '-')

# Hourly data 
  WxData$HourlyWx <- 
    read_xlsx('./data/CGREC_weather.xlsx', 'hourly') %>%
    filter(between(Hour, 1000, 1700)) %>%
    unite( c(Year, Month, Day), col = 'date', sep = '-', remove = F) %>%
    unite( c(Month, Day), col = 'day', sep = '-', remove = F) %>%
    mutate(day = as.Date(day, '%m-%d'), 
           season = case_when(
             between(day, seasons$start[1], seasons$end[1]) ~'Spring', 
             between(day, seasons$start[2], seasons$end[2]) ~'Summer',
             TRUE ~ NA      
           ) ) %>%
    mutate(e = 6.11 * (10 ^ ( (7.5 * DewPoint)/ (237.3 + DewPoint) ) ), 
           es = 6.11 * (10 ^ ( (7.5 * AirTemp)/ (237.3 + AirTemp) ) ), 
           VPD = es - e) 
# Get burn day weather
  WxData$BurnDayWx <-
    HourlyWx  %>% 
    filter(date %in% BurnDays$date) %>%
    group_by(Year, date, season) %>%
    summarise_at(.vars = vars(c(DewPoint, RelHum, WindSpeed,VPD)),
                 .funs = c("mean")) %>%
    ungroup() %>%
    pivot_longer(names_to = 'variable', 
                 values_to = 'value', 
                 cols = DewPoint:VPD) %>%
    mutate(variable = recode(variable, 
                             'DewPoint' = 'Dew Point (°C)', 
                             'RelHum' = 'Relative Humidity (%)', 
                             'VPD' = 'Vapor Pressure Deficit', 
                             'WindSpeed' = 'Wind speed (m/s)') , 
           Year = as.character(Year)) 
\end{lstlisting}