### 2024-12-11 SpringSummerSeverity/paper/CoverLetter/CoverLetter0.tex

\documentclass[parskip=half, 
			   fontsize=12pt,
			   paper=a4]				
{scrartcl}
\input{SetupOptions} % contains scrletter

\newcommand{\journal}{\emph{Restoration Ecology}}

\vspace{-5em}

\begin{document}
	\begin{letter}{Editorial staff\\
		\journal }
\setlength{\parindent}{12pt}

\opening{Dear Sir or Madam,}

Please consider the attached submission for publication in \journal. 
We are excited to leverage a unique dataset to address a novel question about the restoration of pre-European fire regimes: How successful are attempts at summer burning?
Much research attention has been given to variation in fire effects across different seasons, but to our knowledge no one has yet directly addressed the feasability of burning during the growing season, along wth a comparison of burn outcomes (not fire effects) relative to conditions of the fire environment.

We also provide a historical context for the fuel and weather variables considered here, however the limited spatial resolution of the research station represents a narrow scope. 
Indeed, while the relatively high number of 40 fire events across four years provides plenty of food for thought, the present paper is best approached as a case study with a narrow scope in which we demonstrate how various data products can be used to assess patterns in fire activity and outcomes at broader scales.

We've provided suggested reviewers as part of the online submission.

We look forward to working with the editors and reviewers of \journal~to ensure our manuscript meets the standards of the journal.

\vspace{-3em} 
\closing{} %eg. Regards
%\cc{another dude}
\end{letter}

\end{document}