### 2025-04-21 NorthDakotaFireSurvey/R/ResponsePlotting.Rmd

---
title: "Rancher perceptions of Rx fire in North Dakota"
subtitle: "Survey responses"
date: "Version date: `r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: TRUE
    toc_float: FALSE
    theme: journal
    highlight: tango
    code_folding: hide
    code_download: true
    self_contained: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, eval = TRUE, fig.path='../Writing/figures/', dev=c('png'))
# load packages
  pacman::p_load(tidyverse, magrittr, sf, grid, gridExtra, cowplot)
  pacman::p_load_gh('devanmcg/wesanderson')
  ER_pal = wes_palette("Zissou1")[c(1, 4, 5, 2)]
# load pre-wrangled data
  load('./Data/responses.Rdata')
```

# Synthesis and potential recommentations

* Despite not relying on scientific products themselves, respondents do expect administrators of grazinglands to use the best available science
* Respondents generally trust agencies and NGOs, but not necessarily with fire
* Respondents believe their community sees little value in Rx fire and do not want its use expanded
* Initiatives related to Rx fire are likely to be best recieved when they come from Pheasants Forever and Extension
* To increase palatability of Rx fire on public grazinglands, ranchers will likely need to come to appreciate burning as a component of good stewardship, not just a tool for specific management objectives

# Results

## Perceptions of, and trust in, various organizations

Take-aways regarding general use and trust:

* Respondents generally trust most resources
* County Extension resources are the most used
* BLM resources are the least used, consistent with BLM footprint in ND

Take-aways regarding fire, specifically:

* Overall trust is considerably lower with fire
* USFS is most associated with Rx fire, but trusted least
* Pheasants Forever is associated with Rx fire and most trusted with it
* Extension has a high degree of trust but low association

```{r agency_responses, fig.height=12, fig.width=10, fig.cap='Perceptions of various federal (US) and state (ND) agencies and non-governmental organizations regarding (A) use and trust in provided resources, in general; and (B) association and trust with respect to prescribed fire, specifically. TNC = The Nature Conservancy, NRCS = USDA Natural Resource Conservation Service; BLM = US Dept. of Interior Bureau of Land Management; ARS = USDA Agricultural Research Service.'}
# Prepare agency response 
  agency_responses <- 
    responses %>%
      filter(QuestionKey != 'Agency_Resources', 
             str_starts(QuestionKey, "Agency") ) %>%
      mutate(QuestionKey = str_remove(QuestionKey, "Agency_"), 
             QuestionKey = str_replace_all(QuestionKey,  "([[:upper:]])"," \\1"), 
             QuestionKey = str_trim(QuestionKey)) %>% 
      separate(QuestionKey, into = c('category', 'topic'), sep = ' ') %>%
      mutate(agency = case_when(
                        OptionKey == 'NDGF' ~      'ND Game & Fish', 
                        OptionKey == 'PF' ~        'Pheasants Forever', 
                        OptionKey == 'USFWS' ~     'US Fish & Wildlife', 
                        OptionKey == 'audubon' ~   'Audubon Society', 
                        OptionKey == 'extension' ~ 'County Extension', 
                        OptionKey == 'USFS' ~      'US Forest Service', 
                        TRUE ~ str_to_upper(OptionKey) ))
# Resources in general 
  agency_general <- 
    agency_responses %>%
      filter(category != "Fire") %>%
      mutate(Response = recode(topic, 
                               "Trust" = "Trust", 
                               "Use" = "Use"), 
             Response = fct_rev(Response)) %>%
      ggplot() + theme_bw(16) + 
      geom_vline(xintercept = 0, lty = 2) + 
      geom_errorbarh(aes(xmin = min, 
                         xmax = max, 
                         y = agency, 
                         color = Response), 
                     lwd = 0.9,
                     position = position_dodge(width = 0.6)) +
      geom_point(aes(x = mid, y = agency, 
                     fill = Response, 
                     shape = Response), 
                 size = 4, 
                 stroke = 1,
                 position = position_dodge(width = 0.6)) + 
     coord_cartesian(xlim = c(-1.15, 1.9)) + 
      labs(title = 'A. "In general, I personally use/trust resources from this organization"', 
           x = 'Agreement Index') + 
      scale_shape_manual(values = c(21, 24)) + 
      scale_color_manual(values = wes_palette("FantasticFox1")[c(3,5)]) + 
      scale_fill_manual(values = wes_palette("FantasticFox1")[c(3,5)]) + 
      theme(axis.title.y = element_blank(), 
            axis.text.y = element_text(color = 'black', size = 14), 
            plot.title.position = "plot", 
            legend.position = 'top', 
            legend.title=element_blank() )
# WRT fire, specifically
  agency_fire <- 
    agency_responses %>%
      filter(category == "Fire") %>%
      mutate(Response = recode(topic, 
                            "Trust" = "Trust regarding Rx fire", 
                            "Use" = "Associate with Rx fire")) %>%
      ggplot() + theme_bw(16) + 
      geom_vline(xintercept = 0, lty = 2) + 
        geom_errorbarh(aes(xmin = min, 
                           xmax = max, 
                           y = agency, 
                           color = Response), 
                       lwd = 0.9,
                       height = 0.5,
                       position = position_dodge(width = 0.6)) +
        geom_point(aes(x = mid, y = agency, 
                       fill = Response, 
                       shape = Response), 
                   size = 4, 
                   stroke = 1,
                   position = position_dodge(width = 0.6)) + 
       coord_cartesian(xlim = c(-1.15, 1.9)) + 
      labs(title = 'B. "I associate/trust this organization with Rx fire activity or information"', 
           x = 'Agreement Index') + 
       scale_shape_manual(values = c(21, 24)) + 
      scale_color_manual(values = wes_palette("FantasticFox1")[c(3,5)]) + 
      scale_fill_manual(values = wes_palette("FantasticFox1")[c(3,5)]) + 
        theme(axis.title.y = element_blank(), 
              axis.text.y = element_text(color = 'black', size = 14), 
              plot.title.position = "plot", 
              legend.position = 'top', 
              legend.title=element_blank() )
plot_grid(agency_general, 
          agency_fire, 
           # labels = c('A', 'B'),
          ncol = 1)
```

## Contributions to management decisionmaking

Take-aways:

* Respondents mostly value open rangeland 
* Respondents have little concern for managing wildfire risk or invasive species
* Respondents are mostly influenced by a sense of stewardship and passed-down knowledge
* Little direct impact of neighbors and science on decisionmaking

```{r decisionmaking, fig.height=8, fig.width=10, fig.cap='Degree to which respondents agreed that various management considerations (A) were important in their decisions, and (B) influenced their decisionmaking.'}
importance <-
    responses %>%
      filter(str_starts(QuestionKey, "Importance") ) %>%
      mutate(response = case_when(
              OptionKey == 'restore_prairie' ~      'Restore prairie/grassland', 
              OptionKey == 'reduce_wildfire' ~        'Reduce wildfire risk', 
              OptionKey == 'protect_wildlife' ~     'Protect wildlife habitat', 
              OptionKey == 'heterogeneity' ~   'Increase plant diversity', 
              OptionKey == 'control_woody' ~ 'Woody plant control', 
              OptionKey == 'control_invasives' ~      'Invasive species control' )) %>%
  ggplot() + theme_bw(16) + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_errorbarh(aes(xmin = min, 
                     xmax = max, 
                     y = fct_reorder(response, mid)), 
                 color = wes_palette("FantasticFox1")[3],
                  height = 0.5,
                 lwd = 0.9 ) +
  geom_point(aes(x = mid, y = response ),
             fill = wes_palette("FantasticFox1")[3], 
             pch = 21,
             size = 5, 
             stroke = 1 ) + 
  # coord_cartesian(xlim = c(-1.15, 1.9)) + 
  labs(title = 'A. "This environmental concern is important to me"', 
       x = 'Agreement Index') + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_text(color = 'black', size = 14), 
        plot.title.position = "plot" )
influence <- 
    responses %>%
    filter(str_starts(QuestionKey, "Influence") ) %>%
    mutate(response = case_when(
              OptionKey == 'research' ~      'Scientific research/\nUniversity publications', 
              OptionKey == 'neighbors' ~     'How my neighbors\nmanage their land', 
              OptionKey == 'good' ~          'Being a good steward', 
              OptionKey == 'gen' ~   'What previous generations\ntaught me', 
              OptionKey == 'advice' ~ 'Professional advice' )) %>%
    ggplot() + theme_bw(16) + 
    geom_vline(xintercept = 0, lty = 2) + 
    geom_errorbarh(aes(xmin = min, 
                       xmax = max, 
                       y = fct_reorder(response, mid)), 
                   color = wes_palette("FantasticFox1")[5],
                   height = 0.5,
                   lwd = 0.9 ) +
    geom_point(aes(x = mid, y = response ),
               fill = wes_palette("FantasticFox1")[5], 
               pch = 21,
               size = 5, 
               stroke = 1 ) + 
    # coord_cartesian(xlim = c(-1.15, 1.9)) + 
    labs(title = 'B. "This consideration affects my management decisionmaking"', 
         x = 'Agreement Index') + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_text(color = 'black', size = 14), 
          plot.title.position = "plot" )
plot_grid(importance, 
          influence, 
           # labels = c('A', 'B'),
          ncol = 1)
```

## Attitudes about fire

Take-aways

* Respondents generally believe their community has negative perceptions of Rx fire, except for wildlife
* Respondents have a conservative attitude towards Rx fire use and show a disinclination to use it themselves

```{r attitudes, fig.height=10, fig.width=10, fig.cap='Degree to which respondents agreed with statements about (A) community perceptions of prescribed fire effects, and (B) personal attitudes about prescribed fire use.'}
# Attitudes and beliefs about Rx fire
beliefs <-
  responses %>%
    filter(QuestionKey == "CommunityFireBeliefs")  %>%
    mutate(OptionKey = str_remove(OptionKey, 'burn_'), 
           response = case_when(
              OptionKey == 'wildlife' ~ 'Negative impacts to wildlife', 
              OptionKey == 'soil_erosion' ~   'Results in soil erosion', 
              OptionKey == 'smoke' ~    'Produces harmful smoke', 
              OptionKey == 'safety' ~   'Poses risk to human safety', 
              OptionKey == 'property_risk' ~   "Poses risk to nearby property", 
              OptionKey == 'forage' ~      'Results in loss of forage' )) %>%
    ggplot() + theme_bw(16) + 
    geom_vline(xintercept = 0, lty = 2) + 
    geom_errorbarh(aes(xmin = min, 
                       xmax = max, 
                       y = fct_reorder(response, desc(SurveyOrder))) , 
                   color = wes_palette("FantasticFox1")[3], 
                   lwd = 0.9 ) +
    geom_point(aes(x = mid, y = response ),
               fill = wes_palette("FantasticFox1")[3], 
               pch = 21,
               size = 4, 
               stroke = 1 ) + 
    # coord_cartesian(xlim = c(-1.15, 1.9)) + 
    labs(title = 'A. "Members of my community generally believe the following about Rx fire"', 
         x = 'Agreement Index') + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_text(color = 'black', size = 14), 
          plot.title.position = "plot" )
attitudes <-
  responses %>%
    filter(str_starts(QuestionKey, "Attitudes") ) %>%
    mutate(OptionKey = str_remove(OptionKey, 'burn_'), 
           response = case_when(
             OptionKey == 'beneficial' ~ 'Rx fire is a beneficial\ntool for managing rangelands', 
             OptionKey == 'frequent' ~   'People should be able to\nburn as frequently as they\n want for management', 
             OptionKey == 'limited' ~    'Burning can be used in a\nlimited set of instances', 
             OptionKey == 'man_plan' ~   'My management plan includes\nusing prescribed fire', 
             OptionKey == 'prepared' ~   "I'm prepared to use Rx fire\nif I choose to", 
             OptionKey == 'pros' ~      'Fire should only be used\nby professionals', 
             OptionKey == 'shouldnt' ~   'Burning is a poor management\noption that shouldn’t be used', 
             OptionKey == 'training' ~ 'Burning is a good option\nfor anyone with training')) %>%
    ggplot() + theme_bw(16) + 
    geom_vline(xintercept = 0, lty = 2) + 
    geom_errorbarh(aes(xmin = min, 
                       xmax = max, 
                       y = fct_reorder(response, desc(SurveyOrder))) ,
                   height = 0.5,
                   color = wes_palette("FantasticFox1")[5], 
                   lwd = 0.9 ) +
    geom_point(aes(x = mid, y = response ),
               fill = wes_palette("FantasticFox1")[5], 
               pch = 21,
               size = 4, 
               stroke = 1 ) + 
    # coord_cartesian(xlim = c(-1.15, 1.9)) + 
    labs(title = 'B. "I personally believe the following"', 
         x = 'Agreement Index') + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_text(color = 'black', size = 14), 
          plot.title.position = "plot" )
plot_grid(beliefs, 
          attitudes, 
          rel_heights = c(0.75, 1.5), 
          ncol = 1)
```

## Public grazing land management and personal recreation

Take-aways

* Respondents expect administrators to use good science
* Respondents do not want to see Rx fire expanded on public grazingland
* Respondents really don't use public grazlinglands for recreation

```{r public_lands, fig.height=5, fig.width=10, fig.cap='Respondent attitudes towards public grazingland administration and prescribed fire use.'}
responses %>%
  filter(QuestionKey == "PublicGrazingBeliefs", 
         OptionKey != 'public_fire')  %>%
  mutate(response = case_when(
           OptionKey == 'increase_public_fire' ~ 'There should be more Rx fire\non public grazingland',            OptionKey == 'amount_public_fire' ~'There is an appropriate amount of\nRx fire on public grazingland', 
           OptionKey == 'admin_users' ~ 'Administrators should manage to meet\nuser needs or preferences', 
           OptionKey == 'admin_science' ~ "Administrators should manage land\naccording to the best available science", 
           OptionKey == 'admin_good' ~ 'Public grazingland administrators\n do a good job managing the land' )) %>%
  ggplot() + theme_bw(16) + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_errorbarh(aes(xmin = min, 
                     xmax = max, 
                     y = fct_reorder(response, desc(SurveyOrder))) , 
                 color = wes_palette("FantasticFox1")[3], 
                 height = 0.5,
                 lwd = 0.9 ) +
  geom_point(aes(x = mid, y = response ),
             fill = wes_palette("FantasticFox1")[3], 
             pch = 21,
             size = 4, 
             stroke = 1 ) + 
  # coord_cartesian(xlim = c(-1.15, 1.9)) + 
  labs(title = '"I personally believe the following about public grazingland management"', 
       x = 'Agreement Index') + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_text(color = 'black', size = 14), 
        plot.title.position = "plot" )
```

```{r recreation, fig.height=5, fig.width=10, fig.cap="Respondents' use of public grazingland for various recreation activies."}
responses %>%
  filter(QuestionKey == 'Recreation' ) %>%
  ggplot() + theme_bw(16) + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_errorbarh(aes(xmin = min, 
                     xmax = max, 
                     y = fct_reorder(SurveyText, mid) ) , 
                 color = wes_palette("FantasticFox1")[3], 
                 height = 0.5,
                 lwd = 0.9 ) +
  geom_point(aes(x = mid, y = SurveyText ),
             fill = wes_palette("FantasticFox1")[3], 
             pch = 21,
             size = 4, 
             stroke = 1 ) + 
  # coord_cartesian(xlim = c(-1.15, 1.9)) + 
  labs(title = '"Myself and/or my family use public grazingland for these recreation activities"', 
       x = 'Agreement Index') + 
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_text(color = 'black', size = 14), 
        plot.title.position = "plot" )
```