### 2025-04-21 SpringSummerSeverity/paper/CoverLetter/CoverLetterEM1.tex

\documentclass[parskip=half, 
			   fontsize=10pt,
			   paper=a4]				
{scrartcl}
\input{SetupOptions} % contains scrletter

\newcommand{\journal}{\emph{Environmental Management}}

\vspace{-5em}

\begin{document}
	\begin{letter}{Editorial staff\\
		\journal }
\setlength{\parindent}{12pt}

\opening{Dear Sir or Madam,}

Please find attached our revised manuscript for further consideration for publication in \journal. 
We appreciate the opportunity to consider the thoughtful comments of both reviewers. 
The editor will note that we have responded strongly to the suggestions for statistical analysis and the greater precision in discussing results these analyses provide, as well as better definitions of burn objectives, and an expanded discussion of global patterns in phenology as they interact with local factors.

Two points from the reviewers have not been incorporated: Reviewer 1's comments about field data from the fires cannot be satisfied as these data do not exist, and comments from both reviewers about the role of grazing were not incorporated into the paper as we do not have anything substantial to say about a factor we did not observe to have any impact on the variables reported here. 
Certainly, pages can be written about the potential effects of grazing on a fuelbed but without any way to connect grazing to the data we present here, the whole topic strikes us as tangential.

Please recall that, from our initial submission, we have prepared this as a \textbf{Methodology} paper, for two basic reasons: Firstly, our emphasis is on the feasibility of conducting the treatments, not evaluating their ecological outcomes with reference to experimental hypotheses; and secondly, this is a case study focused on a single research station. 
As a reviewer pointed out, although 40+ fires is a decent sample size, we are reticent to extrapolate the conclusions beyond the local context beyond the caveat of the case study.

We look forward to working with the editors and reviewers of \journal~to ensure our manuscript meets the standards of the journal.

\vspace{-3em} 
\closing{} %eg. Regards
%\cc{another dude}
\end{letter}

\end{document}